# Building Maps in R Studio

## Load your libraries and set your working directory

```{r}
# Load libraries
library(tmap)
library(sf)
library(tidyverse)
```

If you don't already have the packages installed, you must install them before loading the libraries; you can install the libraries with the following:

```{r, eval=FALSE}
install.packages(c("tmap", "sf", "tidyverse"))
```

After your libraries are successfully installed and loaded, set your working directory by passing the file path of the directory that contains your data to the ```setwd``` function, as in the example below (but note your file path will look different). 

```{r, eval=F}
# Set your working directory to the directory containing your data
setwd("/Users/adra7980/Documents/git_repositories/opensource_mapping/data")
```

If you don't know the relevant file path, you can set your working directory manually by opening the ```Session``` menu, scrolling down to 

## Read in a spatial dataset
```{r, echo=-1}
setwd("/Users/adra7980/Documents/git_repositories/opensource_mapping/data")
# Read in the shapefile named "usa_shapefile.shp"
usa_shapefile<-st_read("usa_shapefile.shp")
```

